---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
# Clear global env
rm(list = ls())

# Set working directory to data
setwd('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data')
```


# Glucotypes reveal new patterns of glucose dysregulation

Link: https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2005143#sec011

S1 Data. Continuous glucose monitor recordings.
Glucose concentration recorded every 5 minutes. Columns are DisplayTime (time stamp of recording as shown by the monitor, used for analysis), GlucoseValue (glucose concentration in mg/dl), subjectId (participant id), and InternalTime (internal time stamp of recording).

https://doi.org/10.1371/journal.pbio.2005143.s010

This data is from an experiment that measured continuous readings from both healthy and non healthy individuals. In this book I will inspect and preprocess the data. 

The meta data can be viewed here 

```{r}
setwd('C:\\Users\\Harry\\Documents\\Individual project\\data\\Cont_gluc')

library(RSQLite)

filename <- "pbio.2005143.s014.db"
sqlite.driver <- dbDriver("SQLite")
db <- dbConnect(sqlite.driver,
                dbname = filename)
                
## Some operations
dbListTables(db)
mytable <- dbReadTable(db,"clinical")
```


```{r}
View(mytable)
```


## 1. Reading in the data

I read in the data as follows:

```{r}
# Read in data with tab delimination
data <- read.table('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/cont_gluc/pbio.2005143.s010',header = T,sep = '\t')
```

we can then  inspect:

```{r}
# Inspect the data 
View(data)
```

## 2. formatting the data 

The internal time column has the date and time as one string. I want to seperate by time and into individual days, so these will have to be separated into separate columns.

I've written functions to parse the dates and times in the DisplayTime column. These can be found in Helpers.R

```{r}
# Load helper functions
source('Helpers.R')

# Get dates and times from first column
dates_times <-as.array(data[,'InternalTime'])

# Get length for progress indicator
len <- length(dates_times)

# Create arrays for dates and times
dates <-list()
times <- list()

# loop through each entry
for (i in 1:length(dates_times)){
  
  # parse dates and times using helper functions
  new_date <- get_date(dates_times[[i]])
  new_time <- get_time(dates_times[[i]])
  # Add to empty lists 
  dates <- append(dates, new_date)
  times <- append(times, new_time)
}

# Check output 
head(dates)
head(times)
```

Now i can reformat the dates to remove the commas and add to the original data.

```{r}
# loop through dates 
for (i in 1:length(dates)){
  # add to df
  data[i,'date'] <- dates[[i]]
}
```

Same as above for times:

```{r}
# loop through times
for (i in 1:length(dates)){
  # replace commas and add to df
  data[i,'time'] <- times[[i]]
}

View(data)
```

Finally i can remove the redundant columns from the data and save it.

```{r}
# Set columns to remove
rem <- c('DisplayTime', 'InternalTime')

# Filter out columns
final_data <- data[ , !(names(data) %in% rem)]

# Reorder columns with date and time first
final_data <- final_data[,c(3,4,1,2)]


# Set working directory to data
setwd('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/Cont_gluc')

# Save 
write.csv(final_data,file = 'Cont_monit_data.csv')
```

```{r}
# Inspect
View(final_data)
```

Finally I can convert the times into minutes to make it easier for the model to process.

```{r}
library(lubridate)
times2 <- hms(times) 
minutes <- hour(times2)*60 + minute(times2)
final_data['minutes'] <- minutes
```



```{r}
View(final_data)
```


```{r}
# Set working directory to data
setwd('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/Cont_gluc')
write.csv(final_data,file = 'Cont_monit_data.csv')
```

Now im gonna save each day in a separate file to make the data easier to manage and index

```{r}
# Load in data from shared file
data <- read_csv('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/Cont_gluc/Cont_monit_data.csv')

View(data)
```
I can get each unique day as below;

```{r}
# get each unique day 
data$date <-format(data$date, format = "%Y/%m/%d")
days <- unique(data[,'date'])
days <- as.list(days)
```

Now we can filter out each day and save them to a file.

```{r}
# Set working directory
setwd('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/cont_gluc/Days')

for(i in 1:length(days[[1]])){
  day <- days[[1]][i]
  filt <-  filter(data, date == day)
  filename <- paste0(gsub('/','-',day),'.csv') 
  write.csv(filt, file = filename)
  }
```


## Graphing the data 
Its difficult to visualise the data in this format so i generated graphs of each day as follows:

```{r}
files <- list.files(path="/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/cont_gluc/Days", pattern="*.csv", full.names=TRUE, recursive=FALSE)

setwd('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/cont_gluc/Days/Graphs')
for (file in files){
  date <- df$date[1]
  df <- read.csv(file = file)
  x <- df$minutes
  y <- df$GlucoseValue
  ggplot(data = df, aes(x = x, y = y)) + geom_point() + geom_line(aes(group = 1)) + theme(axis.text.x=element_text(color="#993333",size=2, angle=90)) + xlab('Time (minutes)') + ylab('Glucose Value')
  filename <- paste0(gsub('/','-', date),'.png')
  ggsave(filename)
}
```


```{r}
# Set directory
setwd('/Users/harrypetchsmith/Msc Bionformatics/Individual project/data/cont_gluc/Days')
# Load in first day
day1 <- read.csv('2017-07-15.csv')


date <- day1$date[1]
filename <- paste0(gsub('/','-',day),'.pdf') 
date 
```

2015-05-11
-12
-13




