---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
# Predicting blood glucose using a multiple output GP.

The data can be found under: ~/individual project/data/cont_gluc. The data is split up into days. 

## 1. Inspecting the data 
Each data file represents a day of glucose monitoring. First i will visualise some of the data.

The first day (as appears in original data) is as follows
```{r}
# Set directory
setwd('~/Individual project/data/Cont_gluc/Days')
# Load in first day
day1 <- read.csv('2015-05-11.csv')
# Inspect
View(day1)

```

Line plot of day 1:

```{r}
library(ggplot2)

# Get the first 134 elements so both datasets are equal
day1 = day1

# Set x and y values
x = day1$minutes
y = day1$GlucoseValue

# Generate plot of first day and save
plot1 <- ggplot(data = day1, aes(x = x, y = y)) + geom_point(colour = '#0886bd') + ggtitle('2015/05/11 Continous Monitoring Glucose Data') + xlab('Time (minutes)') + ylab('Glucose Value')
ggsave('2015-05-11 Continous Monitoring Glucose Data.pdf')
```
lets check out the subsequent day

```{r}
# Set directory
setwd('~/Individual project/data/Cont_gluc/Days')
# Read day2 file
day2 <- read.csv('2015-05-12.csv')
# inspect
View(day2)
```

We can plot this data as follows

```{r}
library(ggplot2)
# Filter out entries to the same size as day1
day2 <- day2[1:279,]

# set x and y values
x = day2$minutes
y = day2$GlucoseValue

# Generate plot of day2 and save
plot2 <- ggplot(data = day2, aes(x = x, y = y)) + geom_point(colour = 'red') + ggtitle('2015/05/12 Continous Monitoring Glucose Data') + xlab('Time (minutes)') + ylab('Glucose Value')
ggsave('2015-05-12 Continous Monitoring Glucose Data.pdf')
```
I've set these two days to have the same number of values. Day2 was much bigger and required a lot of time to run.

After initial inspection of the data i began to build the model on the data.

## 2. Building the model
The model source can be found here:
Code/R code/2GP_functions.R

```{r}
source('C:/Users/Harry/Documents/Individual project/Code/R code/2GP_functions.R')
```

First i define my time series and Y values for each of the two days.
```{r}
# Time series for day1
t1 <- day1$minutes
#Y vakues for day1
Y1 <- day1$GlucoseValue

#Time series for day2
t2 <- day2$minutes
# Y values for day2
Y2 <- day2$GlucoseValue

# Overall total of time series 
T = day1$minutes
```

The time series defined above were combined into vectors below. 

```{r}
# vector with both time series 
t <- c(t1, t2)
# Vector with Y values
Y <- c(Y1, Y2)
# Y should be 134*2
length(Y)
# number of samples in each series - 134
N1 <- length(t1)
N2 <- length(t2)
```
I generated a combined plot of the two days as follows:

```{r}
# df for day1
day1 <- data.frame(x = t1, y = Y1)
#df for day2
day2 <- data.frame(x = t2, y = Y2)

# Shared plot of both days
plot3 <- ggplot() + 
geom_line(data=day1, aes(x, y,group = 1), colour = '#0886bd') + 
geom_line(data=day2, aes(x, y, group = 1), color='red') + ggtitle('2015/05/11 and 2015/05/12 Continous Monitoring Glucose Data') + xlab('Time (minutes)') + ylab('Glucose Value')
ggsave('shared.pdf') 
```
The data was inspected in the window below 
```{r}
library(ggpubr)
# Generate figure panels
ggarrange(plot1,plot2,plot3,
          nrow = 3,labels = c('A','B','C'))
                  
```
Since this is a Baysian method, we work off a prior assumption and update this based on obsereved data. Our prior assumptions were made as to the distribution of the data and parameters

I am assuming a normal distribution mean and standard deviation for the prior distribution. Parameters were sampled from this distribution:

```{r}
#--------Hyper-parameters--------------------------------------------
# Parametrs = [V,F,W,G,B,mu]
mean_vi <- 0
sd_vi <- 3

mean_wi <- 0
sd_wi <- 3

mean_fi <- 0
sd_fi <- 3

mean_gi <- 0
sd_gi <- 3

mean_bi <- -2
sd_bi <- 3

mean_mu <- 0.5
sd_mu <- 3

# Prior parameters vector
pr_par <- c(mean_vi, sd_vi, mean_wi, sd_wi, mean_fi, sd_fi, mean_gi, sd_gi, mean_bi, sd_bi, mean_mu, sd_mu)
```

These parameters can be sampled randomly from their prior distributions as initial guesses:

```{r}
#--------Sampling--------------------------------------------
vi <- rnorm(2, mean_vi, sd = sd_vi) # A_i = \exp{f_i}, B_i = \exp{g_i}; \sigma_i = \exp{\beta_i}, so I will need to square it later
wi <- rnorm(2, mean_wi, sd = sd_wi)
fi <- rnorm(2, mean_fi, sd = sd_fi)
gi <- rnorm(2, mean_gi, sd = sd_gi)
bi <- rnorm(2, mean_bi, sd = sd_bi)
mu <- rnorm(1, mean_mu, sd = sd_mu)

par <- c(vi[1], wi[1], fi[1], gi[1], bi[1], vi[2], wi[2], fi[2], gi[2], bi[2], mu) #parameters to estimate; [v1 w1 f1 g1 Beta1 v2 w2 f2 g2 Beta2 mu];
```

Lets try calculating a covariance matrix from these parameters. This was doen as an example to test if the code works.

```{r}
# Calculate covariance matrix
D <- Covariance(par, t1, t2)
# View covariance matrix as a data frame 
df <- data.frame(matrix(unlist(D), nrow=length(D), byrow=TRUE))
View(D)
```

Now i train the model.

```{r}
source('C:/Users/Harry/Documents/Individual project/Code/R code/2GP_functions.R')
par_optim <- optim(par, log_lik, method = "Nelder-Mead", Y = Y, t1 = t1, t2 = t2, prior = pr_par)
```

once the model is trained i retrieve the optimal parameters

```{r}
# Optimised parameters
par_opt <- par_optim$par

# Calculte covarinace matrix
C <- Covariance(par_opt, t1, t2)

# Solve matrix for smaller values in calculations
C_I <- solve(C)

# Get samples for time series
s_in <- seq(min(T), max(T), length = length(t1))
```

Finally we can make the calculations to retrieve our posterior outputs:

```{r}
#----------------OUTPUT1_Y1------------------------------------------------------------------------
#Cross covariance
PR12 <- c(par_opt[1], par_opt[6], par_opt[3], par_opt[8], par_opt[11]) #v1 v2 f1 f2 mu
# Autocovariance
PR11 <- par_opt[1:4] # v1,w1,f1,g1
kapa1 <- C_ii(PR11, 0) + exp(par_opt[5])^2   # or kapa1 = par_opt(1)^2+par_opt(2)^2+exp(par_opt(5))^2   #v_1^2 + w_1^2 + sigma_1^2
# Empyt vector to store output
output1 <- c()
sg_sqr1 <- c()

# Loop through each sample
for (j in 1:length(s_in)){
	k1 <- c()
	# Calculate K values within and between days
        for (i in 1:N1){
            k1 <- c(k1, C_ii(PR11, s_in[j] - t1[i]))
        }
        for (ii in 1:N2){
            k1 <- c(k1, C_12(PR12, s_in[j] - t2[ii]))
        }
	# Store ouputs and error margins 
        output1 <- c(output1, t(k1) %*% C_I %*% Y)
        sg_sqr1 <- c(sg_sqr1, kapa1 - t(k1) %*% C_I %*% k1)
}

# Upper and lower bounds
y1.low  = output1 - 2 * sqrt(sg_sqr1)
y1.high = output1 + 2 * sqrt(sg_sqr1)
```

The same was performed for the second day 

```{r}
#----------------OUTPUT2_Y2-------------------------------------------------------------------------
#Autocovariance (Cross is same as abpve)
PR22 <- par_opt[6:9] # v2,w2,f2,g2
kapa2 <- C_ii(par_opt[6:9], 0) + exp(par_opt[10])^2 # or kapa2 = par_opt(6)^2+par_opt(7)^2+exp(par_opt(10))^2;

# Empyt vectors for outputs
output2 <- c()
sg_sqr2 <-c()

# Loop thrpugh each sample 
for (j in 1:length(s_in)){
    k2 <- c()
    # Calculate K values 
    for (i in 1:N1){
        k2 <- c(k2, C_21(PR12, s_in[j] - t1[i]))
    }
    for (ii in 1:N2){
        k2 <- c(k2, C_ii(PR22, s_in[j] - t2[ii]))
    }
    # Store ouptus and error
    output2 <- c(output2, t(k2) %*% C_I %*% Y)
    sg_sqr2 <- c(sg_sqr2, kapa2 - t(k2) %*% C_I %*% k2)
}
# Upper and lower bounds
y2.low  = output2 - 2 * sqrt(sg_sqr2)
y2.high = output2 + 2 * sqrt(sg_sqr2)
```

The results were visualised as follows


```{r}
library(ggplot2)

# Visulaise Mean function and error regions.
ggplot(data = data.frame(s_in,output1), aes(x = s_in, y = output1)) + 
  geom_ribbon(aes(ymin = y1.low, ymax = y1.high), colour = '#9BA3F0', fill = '#9BA3F0', alpha = 0.6, size = 0.1) +
  geom_point(data = data.frame(t1,Y1), aes(x = t1, Y1), size = 0.4, colour = 'red') +
  geom_line(data = data.frame(s_in,output1), aes(x = s_in, y = output1), colour = 'green', alpha = 0.8, size = 0.4) + xlab('Time (Minutes)') + ylab('Glucose Value')
```

```{r}
library(ggplot2)

# Visualise mean function and error regions
ggplot(data = data.frame(s_in,output2), aes(x = s_in, y = output2)) + 
  geom_ribbon(aes(ymin = y2.low, ymax = y2.high), colour = '#ED829C', fill = '#ED829C', alpha = 0.6, size = 0.1) +
  geom_point(data = data.frame(t2,Y2), aes(x = t2, Y2), size = 0.4, colour = 'red') +
  geom_line(data = data.frame(s_in,output2), aes(x = s_in, y = output2), colour = 'blue', alpha = 0.8, size = 0.4) + xlab('Time (Minutes)') + ylab('Glucose Value') 
```

