---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Single ouptput GP
In order to compare the efficacy of the mutliple output Gaussian process i am comparing with the standard single output model. To do this i used Gaupro - a library that implements the gaussian process with the Gaussian kernel function.

```{r}
library(GauPro)
```

I insoected the data as follows
```{r}
# Set directory
setwd('~/Individual project/data/Cont_gluc/Days')
# Load in first day
day1 <- read.csv('2015-05-11.csv')
# Inspect
View(day1)
```

```{r}
library(ggplot2)

day1 = day1

# Set x and y values 
x = day1$minutes
y = day1$GlucoseValue

# Visualize 
ggplot(data = day1, aes(x = x, y = y)) + geom_point(colour = 'purple') + theme(axis.text.x=element_text(color="#993333",size=2, angle=90)) + ggtitle('2017-07-15 Continous monitoring glucose data') + xlab('Time (minutes)') + ylab('Glucose Value')
```
The same was done for day2

```{r}
# Set directory to data 
setwd('~/Individual project/data/Cont_gluc/Days')

# Read day 2 file 
day2 <- read.csv('2015-05-12.csv')

# Inspect
View(day2)
```
```{r}
library(ggplot2)
# Get 279 entries of day 2 so sizes are equal
day2 <- day2[1:279,]

# Set x and y values
x = day2$minutes
y = day2$GlucoseValue

# Visualise 
ggplot(data = day2, aes(x = x, y = y)) + geom_point(colour = 'red') + theme(axis.text.x=element_text(color="#993333",size=8, angle=90)) + ggtitle('2017-07-16 Continous monitoring glucose data') + xlab('Time (minutes)') + ylab('Glucose Value')
```
Next I built and ran the model on the data.
As with the multiple output model I first defined the time series and y values for each dataset to insert into the model.

```{r}
# Time series for day1
t1 <- day1$minutes
#Y vakues for day1
y1 <- day1$GlucoseValue

#Time series for day2
t2 <- day2$minutes
# Y values for day2
y2 <- day2$GlucoseValue

# Overall total of time series 
T = day1$minutes
```


```{r}
# --- DAY 1 -----------------------------------------------------------------------------------------
library(GauPro)

# Set x and y values
x = day1$minutes
y = day1$GlucoseValue

# Define the GP with x and y values 
gp <- GauPro(x, y, parallel=FALSE)

# Visualise the model
ggplot(data = data.frame(t1,y1), aes(t1, y1)) + 
  geom_ribbon(aes(ymin = gp$predict(x)+2*gp$predict(x, se=T)$se, ymax = gp$predict(x)-2*gp$predict(x, se=T)$se), colour = '#558587', fill = '#558587', alpha = 0.6, size = 0.1) +
  geom_point(data = data.frame(t1,Y1), aes(x = t1, Y1), size = 1.2, colour = '#0886bd') +
  geom_line(data.frame(x,gp$predict(x)),mapping = aes(x, y=gp$predict(x) ),colour = '#0886bd', size = 1)  + xlab('Time (Minutes)') + ylab('Glucose Value') + ggtitle('2017-07-15 Posterior Output') + theme(panel.border = element_rect(colour = "#381fab", fill=NA, size=2))
```
The same was done for Day 2

```{r}
# Reassign x and y values
x = day2$minutes
y = day2$GlucoseValue

library(GauPro)

# Define the model with x and y
gp <- GauPro(x, y, parallel=FALSE)

# Visualise the model
ggplot(data = data.frame(t2,y2), aes(t2, y2)) + 
  geom_ribbon(aes(ymin = gp$predict(x)+2*gp$predict(x, se=T)$se, ymax = gp$predict(x)-2*gp$predict(x, se=T)$se), colour = '#ED829C', fill = '#ED829C', alpha = 0.6, size = 0.1) +
  geom_point(data = data.frame(t2,Y2), aes(x = t2, Y2), size = 1.2, colour = 'red') +
  geom_line(data.frame(x,gp$predict(x)),mapping = aes(x, y=gp$predict(x) ),colour = 'blue', size = 1)  + xlab('X') + ylab('Sin(2X)') + ggtitle('') + theme(panel.border = element_rect(colour = "#381fab", fill=NA, size=2))
```





